---
title: "centerlines"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(terra)
library(dplyr)
library(sf)
library(glue)
library(fs)
library(stringr)
library(stars)
library(scico)
library(ggplot2)
library(whitebox)
```

# river mosaic to smooth polygons

```{r}
data_dir <- 'S:/users/khondula/DATA'
checked_site_yr <- fs::dir_ls(glue('{data_dir}/centerlines/03_rivmosaic-imgs')) %>% 
  basename() %>% grep('ORNL', x = ., value = TRUE)
checked_site_yr

basename(fs::dir_ls(glue('{data_dir}/centerlines/03_watermask-rivgroups'), glob = glue('*{checked_site_yr}*MNDWI_01.tif')))
```

```{r}
my_group_id <- 'ornl2'
```

# select river segment polygons

# mask mndwi with smoothed river

```{r}
my_riv_fill10 <- st_read(glue('{data_dir}/centerlines/05_riv_smooth_polygons/{checked_site_yr}-{my_group_id}-riverfill10.shp'))
```

```{r}
# rm(my_awei)
# rm(my_awei_max)
my_mndwi_path <- glue('{data_dir}/centerlines/03_watermask-rivgroups/{checked_site_yr}-{my_group_id}-MNDWI_01.tif')
my_mndwi <- terra::rast(my_mndwi_path)
plot(my_mndwi)
```

```{r}
riv_sv <- as(my_riv_fill10, "SpatVector")
# my_mndwi_corp <- terra::crop(my_mndwi, riv_sv)
# ggplot() +
#   geom_stars(data = st_as_stars(my_mndwi_corp)) +
#   geom_sf(data = my_riv_fill10[1,], fill = NA, col = 'red')
# 
# rm3 <- terra::mask(my_mndwi_corp, mask = riv_sv, updatevalue = 0) # make ALL areas inside mask 1
# rm4 <- terra::mask(rm3, mask = riv_sv, updatevalue = 1, inverse = TRUE) # make ALL areas inside mask 1
# plot(rm4)
# gg1 <- ggplot() + geom_sf(data = my_riv_fill10)
# ggsave('testriv.pdf', gg1)
# riv_mask2 <- terra::mask(my_mndwi, mask = riv_sv, inverse = TRUE, updatevalue = 0) # make ALL areas outside mask 0
# plot(riv_mask2)
# riv_mask3 <- terra::mask(my_mndwi, mask = riv_sv, inverse = FALSE, updatevalue = 1) # make ALL areas inside mask 1
riv_mask3 <- terra::mask(my_mndwi, mask = riv_sv, updatevalue = 0) # make ALL areas inside mask 1
riv_mask4 <- terra::mask(riv_mask3, mask = riv_sv, updatevalue = 1, inverse = TRUE) # make ALL areas inside mask 1

plot(my_mndwi)
plot(riv_mask4)
maskpath <- glue('{data_dir}/centerlines/03_watermask-rivgroups/{checked_site_yr}-{my_group_id}-MNDWI_01mask.tif')
riv_mask4 %>% terra::writeRaster(filename = maskpath, overwrite = TRUE)
# zoom(riv_mask3)
```

Distance calculation with whitebox

```{r}
rcl_matrix <- matrix(c(0, 1, 1, 0), ncol = 2, byrow = TRUE)
riv_mask_wb <- terra::classify(riv_mask3, rcl = rcl_matrix)

plot(riv_mask_wb)
zoom(riv_mask_wb)
# save locally for whitebox calculation

wb_dir <- glue('06_rivmask_4wb/{checked_site_yr}')
fs::dir_create(wb_dir)

wb_filename1 <- glue('{wb_dir}/rivmask_wb-{my_group_id}.tif')
riv_mask_wb %>% terra::writeRaster(wb_filename1, overwrite = TRUE)
```

# whitebox distance calculation

```{r}
wb_dir2 <- glue('07_wb-dist/{checked_site_yr}')
fs::dir_create(wb_dir2)
wb_filename2 <- glue('{wb_dir2}/wb_dists-{my_group_id}.tif')

# wbt_euclidean_distance(
#   input = wb_filename1,
#   output = wb_filename2,
#   compress_rasters = TRUE,
#   verbose_mode = TRUE)

riv_dist <- terra::rast(wb_filename2)
plot(riv_dist)
# riv_dist_mask <- terra::mask(riv_dist, riv_smooth_sv, inverse = TRUE)
# zoom(riv_dist)
# plot(riv_dist_mask)
# wb_dir2c <- glue('{data_dir}/centerlines/07_wb-dist/{basename(my_site_dir)}')
# fs::dir_create(wb_dir2c)
# riv_dist %>% terra::wrap() %>% saveRDS(glue('{wb_dir2c}/wb_dists.rds'))
# riv_dist %>% terra::writeRaster(glue('{wb_dir2c}/wb_dists.tif'))

# plot(riv_dist)
```

# try canny edges

```{r}
# library(imager)
# my_riv <- raster::raster('../SWE/NEON_D03_JERC_DP3_738000_3463000_reflectance-rivmask_wb.tif')
# my_riv <- raster::reclassify(my_riv, matrix(c(0,1,1,0), nrow = 2))
# plot(my_riv)
# my_riv %>% writeRaster('testriver.tif')
# riv_cimg <- as.cimg(my_riv)
# riv_px <- riv_cimg > 0
# shrink(riv_px, 10, boundary = FALSE) %>% plot()
# grow(riv_px, 3) %>% plot()
# # riv_canny <- cannyEdges(riv_px)
# riv_canny <- cannyEdges(riv_cimg) 
# riv_canny_cimg <- as.cimg(riv_canny)
# plot(riv_canny_cimg)
# 
# 
# class(riv_canny_cimg)
# st_as_stars(riv_canny_cimg)
# # save.image(riv_canny_cimg, 'canny.tiff')
# # riv_canny_r <- as(riv_canny_cimg, 'Raster')
# class(riv_canny_r)
# # riv_canny_r
# plot(riv_canny_r)
# riv_canny_rast <- terra::rast(riv_canny_r)
# 
# plot(boats)
# cannyEdges(boats, alpha = 0.4) %>% plot()
# 
# im <- imfill(10,10)
# px <- px.all(im)
# plot(im)
# shrink(px,3,bound=TRUE) %>% plot(main="Boundary conditions: TRUE")
# shrink(px,3,bound=FALSE) %>% plot(main="Boundary conditions: FALSE")

```

# edge detection thresholds

```{r}

w_sobel1 <- matrix(c(-1, -2, -1, 0, 0, 0, 1, 2, 1), nrow = 3)
w_sobel2 <- matrix(c(1, 2, 1, 0, 0, 0, -1, -2, -1), byrow = TRUE, nrow = 3)

riv_dists_sobel1 = terra::focal(riv_dist, 
                   w = w_sobel1, 
                   expand = TRUE)
riv_dists_sobel2 = terra::focal(riv_dist, 
                   w = w_sobel2, 
                   expand = TRUE)

riv_sobel <- sqrt(riv_dists_sobel1^2 + riv_dists_sobel2^2)
plot(riv_sobel)
# zoom(riv_sobel)
```


```{r}
# remask with river so areas outside are NA
riv_sobel_mask <- terra::mask(riv_sobel, mask = riv_sv, touches = TRUE)
# plot(riv_sobel_mask)
# plot(riv_sobel)
# zoom(riv_sobel_mask)

# riv_sobel_mask <- terra::mask(riv_sobel, mask = riv_smooth_sv, touches = TRUE)
# riv_sobel_mask2 <- terra::classify(riv_sobel_mask, matrix(c(NA, minmax(riv_sobel_mask)[2]), nrow = 1))

# plot(riv_sobel_mask2)
# zoom(riv_sobel_mask2)
# create points to manually identify river segments
# low threshold for zero crossings
my_val <- quantile(terra::values(riv_sobel_mask), 0.01, na.rm = TRUE) # 0.2 for area, 0.01 for points? 

rcl_matrix_pts <- matrix(c(-Inf, my_val, 1, 
                         my_val, Inf, NA), ncol = 3, byrow = TRUE)

riv_sobel_mask01 <- terra::classify(riv_sobel_mask, rcl = rcl_matrix_pts, include.lowest = FALSE)

# # now higher threshold for area
# my_val2 <- quantile(terra::values(riv_sobel), 0.25, na.rm = TRUE)
my_val2 <- quantile(terra::values(riv_sobel_mask), 0.5, na.rm = TRUE)
my_val2
plot(riv_sobel)
rcl_matrix_area <- matrix(c(my_val2, Inf, NA), ncol = 3, byrow = TRUE)
riv_sobel_mask20 <- terra::classify(riv_sobel_mask, rcl = rcl_matrix_area, include.lowest = FALSE)

plot(riv_sobel_mask20)
zoom(riv_sobel_mask20)
```

```{r}
# convert to points and save
riv_cent_sv <- terra::as.points(riv_sobel_mask01)
riv_cent_sf <- riv_cent_sv %>% st_as_sf() %>% tibble::rowid_to_column('pt_id') %>% dplyr::mutate(rivgroup = my_group_id)
```

```{r}
library(leaflet)
pts4326 <- st_transform(riv_cent_sf, 'EPSG:4326')
riv4326 <- my_riv_fill10 %>% st_transform('EPSG:4326')

leaflet() %>%
  addTiles(group = 'bg') %>%
  addProviderTiles(providers$Esri.WorldImagery, group = 'esri') %>%
  addPolygons(data = riv4326, color = 'cyan', group = 'riv',
              fillOpacity = 1, opacity = 1, label = ~glue('{rand_id}')) %>%
   addMarkers(data = pts4326,   
             clusterOptions = markerClusterOptions(),
             group = 'pts', popup = ~glue('point {pt_id}')) %>%
  addLayersControl(overlayGroups = c('bg','riv', 'esri', 'pts'))
```

```{r}
my_pt_ids <- c(3258, 16457, 16510, 17163)

my_pts_sf <- riv_cent_sf %>% dplyr::filter(pt_id %in% my_pt_ids)
```

LEAST COST PATH

```{r}
riv_sobel %>% writeRaster(glue('{data_dir}/centerlines/shen-sobel.tif'))
my_pts_sf %>% st_write(glue('{data_dir}/centerlines/shen-pts.shp'))
```

```{r}
riv_sobel_r <- gdistance::raster(riv_sobel_mask20)
tr <- gdistance::transition(riv_sobel_r, function(x) 1/mean(x), directions = 8)
```

```{r}
# pt1 <- my_pts_sf %>% dplyr::filter(pt_id == 3258)
# pt2 <- my_pts_sf %>% dplyr::filter(pt_id == 16457)
pt1 <- my_pts_sf %>% dplyr::filter(pt_id == 16510)
pt2 <- my_pts_sf %>% dplyr::filter(pt_id == 17163)
# plot(my_riv_fill10$geometry)
# plot(pt1, add = TRUE, bg = "green")
# plot(pt2, add = TRUE, col - 'red')
# least cost path from points - For each line segment
my_spath_sp <- gdistance::shortestPath(tr, origin = st_coordinates(pt1), goal = st_coordinates(pt2), output = 'SpatialLines')
line_sf <- my_spath_sp %>% st_as_sf()
ggplot() + geom_sf(data = line_sf)
```

```{r}
library(leaflet)
line4326 <- st_transform(line_sf, 'EPSG:4326')

leaflet() %>%
  addTiles(group = 'bg') %>%
  addProviderTiles(providers$Esri.WorldImagery, group = 'esri') %>%
  addPolygons(data = riv4326, color = 'cyan', group = 'riv',
              fillOpacity = 1, opacity = 1, label = ~glue('{rand_id}')) %>%
  addPolylines(data = line4326, col = 'red', opacity = 1, group = 'line') %>%
  addLayersControl(overlayGroups = c('bg','riv', 'esri', 'line'))
```

save line

```{r}
line_sf %>% st_write('shen-riverline2.shp')
```


## end!


old below here

```{r}
# Save points
# plot(riv_cent_sf)
pts_dir <- glue('{data_dir}/centerlines/09_centerpts/{basename(my_site_dir)}')
fs::dir_create(pts_dir)
riv_cent_sf %>% st_write(glue('{pts_dir}/{basename(my_site_dir)}_{my_group_id}-mask01pts.shp'), append = FALSE)
```

```{r}
my_dir8 <- glue('{data_dir}/centerlines/08_sobelmask20/{basename(my_site_dir)}')
fs::dir_create(my_dir8)

wb_filename1 <- glue('{my_dir8}/sobelmask20-{my_group_id}.tif')
riv_sobel_mask20 %>% terra::writeRaster(wb_filename1, overwrite = TRUE)

my_dir9 <- glue('09_wb_patches/{basename(my_site_dir)}')
fs::dir_create(my_dir9)
```

# identify patches

```{r}
# library(Polychrome)
# library(Polychrome)
# kelly.colors(22) %>% scales::show_col()
# createPalette(100,  c("#ff0000", "#00ff00", "#0000ff")) %>% scales::show_col()

# terraOptions()
# terra::mem_info(riv_sobel_mask20)
# plot(riv_sobel_mask20)
# ?patches
patches_filesname <- glue('{my_dir9}/patches-{my_group_id}.tif')
riv_sobel_mask20_patches <- terra::patches(riv_sobel_mask20, directions = 8)

plot(riv_sobel_mask20_patches)

wb_dir9c <- glue('{data_dir}/centerlines/09_patches/{basename(my_site_dir)}')
fs::dir_create(wb_dir9c)
riv_sobel_mask20_patches %>% terra::writeRaster(glue('{wb_dir9c}/patches-{my_group_id}.tif'))

```

maybe need to do on dell 

```{r}
gg3 <- ggplot() +
  geom_sf(data = riv_smooth, col = 'red', fill = NA, lwd = 0) +
  geom_stars(data = st_as_stars(riv_sobel_mask20_patches)) +
  scale_fill_viridis_c() +
  theme_void() +
  theme(legend.position = 'none')
#
ggsave(glue('{wb_dir9c}/{basename(my_site_dir)}-patches-{my_group_id}.pdf'), gg3)

```

# save leaflet

maybe try splitting odd and even patches into separate layers? 

```{r}
is.even <- function(x) x %% 2 == 0
max_val <- minmax(riv_sobel_mask20_patches)[2]
evens_rast <- terra::app(riv_sobel_mask20_patches, is.even)
rcl_matrix <- matrix(c(0, max_val), ncol = 2, byrow = TRUE)
max_rast <- terra::classify(evens_rast, rcl_matrix)
patches2 <- riv_sobel_mask20_patches + max_rast
```

```{r}
gg3 <- ggplot() +
  geom_sf(data = riv_smooth, col = 'red', fill = NA, lwd = 0) +
  geom_stars(data = st_as_stars(patches2)) +
  scale_fill_viridis_c() +
  theme_void() +
  theme(legend.position = 'none')
#
# ggsave(glue('{wb_dir9c}/{basename(my_site_dir)}-patches-{my_group_id}b.jpeg'), gg3)
ggsave(glue('{wb_dir9c}/{basename(my_site_dir)}-patches-{my_group_id}b.pdf'), gg3)
patches2 %>% writeRaster(glue('{wb_dir9c}/{basename(my_site_dir)}-patches-{my_group_id}b.tif'))
```

 
```{r}
library(leaflet)
library(htmltools)

patches4326 <- terra::project(patches2, 'EPSG:4326')
# patches4326 <- terra::project(riv_sobel_mask20_patches, 'EPSG:4326')
# minmax(patches4326)
patches_r <- raster::raster(patches4326)

my_pal <- colorNumeric('viridis', na.color = "#00000000", domain = 1:minmax(patches4326)[2])
# my_pal <- createPalette(minmax(patches4326)[2],  c("#ff0000", "#00ff00", "#0000ff"))


ll <- leaflet(height = 1000) %>%
  addTiles(group = 'bg') %>%
  addProviderTiles(providers$Esri.WorldImagery, group = 'esri') %>%
  addRasterImage(x = patches_r, colors = my_pal, maxBytes = 4e6, group = 'patches') %>%
  addMarkers(data = st_transform(riv_cent_sf, 4326),   
             clusterOptions = markerClusterOptions(),
             group = 'pts', popup = ~glue('point {pt_id}')) %>%
  addLayersControl(overlayGroups = c('bg','pts', 'esri', 'patches'))
# ll
ll %>% htmltools::save_html(file = glue('10_leaflets/leaflet-{basename(my_site_dir)}-{my_group_id}b.html'))
```


