---
title: "centerlines"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(terra)
library(dplyr)
library(sf)
library(glue)
library(fs)
library(stringr)
library(stars)
library(scico)
library(ggplot2)
library(whitebox)
```

# river mosaic to smooth polygons

```{r}
data_dir <- '/Volumes/AOPScratch/users/khondula/DATA'
checked_site_yr <- fs::dir_ls(glue('{data_dir}/centerlines/03_rivmosaic-imgs')) %>% 
  basename() %>% grep('TOOL', x = ., value = TRUE)
checked_site_yr

basename(fs::dir_ls(glue('{data_dir}/centerlines/03_watermask-rivgroups'), glob = glue('*{checked_site_yr}*MNDWI_01.tif')))
```

```{r}
my_group_id <- 'tool3'
```

# select river segment polygons

# mask mndwi with smoothed river

```{r}
my_riv_fill10 <- st_read(glue('{data_dir}/centerlines/05_riv_smooth_polygons/{checked_site_yr}-{my_group_id}-riverfill10.shp'))
```

```{r}
# rm(my_awei)
# rm(my_awei_max)
my_mndwi_path <- glue('{data_dir}/centerlines/03_watermask-rivgroups/{checked_site_yr}-{my_group_id}-MNDWI_01.tif')
my_mndwi <- terra::rast(my_mndwi_path)
plot(my_mndwi)
```

Try to mask -- might need to do on dell

```{r}
# plot(my_riv_fill10$geometry)
riv_sv <- as(my_riv_fill10, "SpatVector")
```

Mask on mac
```{r}
# rm3 <- terra::mask(my_mndwi, mask = riv_sv, updatevalue = 0) # make ALL areas outside mask 0
# plot(rm3)
# riv_mask4 <- terra::mask(rm3, mask = riv_sv, updatevalue = 1, inverse = TRUE) # make ALL areas inside mask 1
# plot(riv_mask4)
# 

```
 OR MASK ON DELL and read in 
```{r}
my_mask_path <- glue('{data_dir}/centerlines/03_watermask-rivgroups/{checked_site_yr}-{my_group_id}-MNDWI_01mask.tif')
riv_mask4 <- terra::rast(my_mask_path)
plot(riv_mask4)
```

Distance calculation with whitebox

```{r}
rcl_matrix <- matrix(c(0, 1, 1, 0), ncol = 2, byrow = TRUE)
riv_mask_wb <- terra::classify(riv_mask4, rcl = rcl_matrix)

plot(riv_mask_wb)
# zoom(riv_mask_wb)
# save locally for whitebox calculation

wb_dir <- glue('06_rivmask_4wb/{checked_site_yr}')
fs::dir_create(wb_dir)

wb_filename1 <- glue('{wb_dir}/rivmask_wb-{my_group_id}.tif')
riv_mask_wb %>% terra::writeRaster(wb_filename1, overwrite = TRUE)
```

# whitebox distance calculation

```{r}
wb_dir2 <- glue('07_wb-dist/{checked_site_yr}')
fs::dir_create(wb_dir2)
wb_filename2 <- glue('{wb_dir2}/wb_dists-{my_group_id}.tif')

wbt_euclidean_distance(
  input = wb_filename1,
  output = wb_filename2,
  compress_rasters = TRUE,
  verbose_mode = TRUE)

riv_dist <- terra::rast(wb_filename2)
plot(riv_dist)
```

# edge detection

```{r}

w_sobel1 <- matrix(c(-1, -2, -1, 0, 0, 0, 1, 2, 1), nrow = 3)
w_sobel2 <- matrix(c(1, 2, 1, 0, 0, 0, -1, -2, -1), byrow = TRUE, nrow = 3)

riv_dists_sobel1 = terra::focal(riv_dist, 
                   w = w_sobel1, 
                   expand = TRUE)
riv_dists_sobel2 = terra::focal(riv_dist, 
                   w = w_sobel2, 
                   expand = TRUE)

riv_sobel <- sqrt(riv_dists_sobel1^2 + riv_dists_sobel2^2)
# plot(riv_sobel)
# zoom(riv_sobel)
```


```{r}
# remask with river so areas outside are NA
riv_sobel_mask <- terra::mask(riv_sobel, mask = riv_sv, touches = TRUE)

# create points to manually identify river segments
# low threshold for zero crossings
my_val <- quantile(terra::values(riv_sobel_mask), 0.01, na.rm = TRUE) # 0.2 for area, 0.01 for points? 

rcl_matrix_pts <- matrix(c(-Inf, my_val, 1, 
                         my_val, Inf, NA), ncol = 3, byrow = TRUE)

riv_sobel_mask01 <- terra::classify(riv_sobel_mask, rcl = rcl_matrix_pts, include.lowest = FALSE)
```

```{r}
# convert to points and save
riv_cent_sv <- terra::as.points(riv_sobel_mask01)
riv_cent_sf <- riv_cent_sv %>% st_as_sf() %>% tibble::rowid_to_column('pt_id') %>% dplyr::mutate(rivgroup = my_group_id)
```


```{r}
sobel_dir <- glue('{data_dir}/centerlines/08_sobel/{checked_site_yr}')
fs::dir_create(sobel_dir)
riv_sobel %>% writeRaster(glue('{sobel_dir}/{checked_site_yr}-{my_group_id}-sobel.tif'), overwrite = TRUE)
pts_dir <- glue('{data_dir}/centerlines/09_centerpts/{checked_site_yr}')
fs::dir_create(pts_dir)
riv_cent_sf %>% st_write(glue('{pts_dir}/{checked_site_yr}-{my_group_id}-pts_all.shp'), append = FALSE)
```

Find end points for each polygon/line segment

```{r}
library(leaflet)
pts4326 <- st_transform(riv_cent_sf, 'EPSG:4326')
riv4326 <- my_riv_fill10 %>% st_transform('EPSG:4326') %>% tibble::rowid_to_column()
my_pal <- colorNumeric('viridis', domain = 1:nrow(riv4326))

leaflet() %>%
  addTiles(group = 'bg') %>%
  addProviderTiles(providers$Esri.WorldImagery, group = 'esri') %>%
  addPolygons(data = riv4326, color = ~my_pal(rowid), group = 'riv',
              fillOpacity = 0.5, opacity = 1, label = ~glue('{rand_id}')) %>%
   addMarkers(data = pts4326,   
             clusterOptions = markerClusterOptions(),
             group = 'pts', popup = ~glue('point {pt_id}')) %>%
  addLayersControl(overlayGroups = c('bg','riv', 'esri', 'pts'))
```

```{r}
all_lines_df <- readxl::read_excel('line_segment_info.xlsx')
all_lines_df %>% readr::write_csv('ref/line_segment_info.csv')
lines_df <- all_lines_df %>%
  dplyr::filter(site_yr_id %in% checked_site_yr, river_group_id %in% my_group_id)
lines_df
my_seg_ids <- lines_df[['seg_id']]
my_seg_ids
```

```{r}
my_pt_ids <- unique(c(lines_df[['pt1']], lines_df[['pt2']]))
my_pts_sf <- riv_cent_sf %>% dplyr::filter(pt_id %in% my_pt_ids)
```

```{r}
pts_dir <- glue('{data_dir}/centerlines/09_centerpts/{checked_site_yr}')
fs::dir_create(pts_dir)
my_pts_sf %>% st_write(glue('{pts_dir}/{checked_site_yr}-{my_group_id}-pts.shp'), append = FALSE)
```

LEAST COST PATH

- ornl2 raster too big for making transition object

```{r}
riv_sobel_r <- gdistance::raster(riv_sobel_mask)
tr <- gdistance::transition(riv_sobel_r, function(x) 1/mean(x), directions = 8)
# tr %>% saveRDS(glue('{data_dir}/centerlines/{checked_site_yr}-{my_group_id}-tr-all.rds'))
```

# make line segment

```{r}
# my_seg_id <- 175

save_spath_line <- function(my_seg_id){
  seg_df <- dplyr::filter(lines_df, seg_id %in% my_seg_id)
  pt1 <- my_pts_sf %>% dplyr::filter(pt_id %in% seg_df[['pt1']])
  pt2 <- my_pts_sf %>% dplyr::filter(pt_id %in% seg_df[['pt2']])
  # least cost path from points - For each line segment
  my_spath_sp <- gdistance::shortestPath(tr, origin = st_coordinates(pt1), goal = st_coordinates(pt2), output = 'SpatialLines')
  line_sf <- my_spath_sp %>% st_as_sf() %>% dplyr::mutate(seg_id = my_seg_id, site_yr_id = checked_site_yr, group_id = my_group_id)
  gg1 <- ggplot() + geom_sf(data = my_riv_fill10, fill = 'dodgerblue', lwd = 0) + geom_sf(data = line_sf)
  
  lines_dir <- glue('{data_dir}/centerlines/10_centerlines/{checked_site_yr}')
  fs::dir_create(lines_dir)
  line_sf %>% st_write(glue('{lines_dir}/{checked_site_yr}-{my_group_id}-riverline-{my_seg_id}.shp'), append = FALSE)
  return(gg1)
}
```

```{r}
my_seg_ids %>% purrr::walk(~save_spath_line(.x))
```

```{r}
lines_dir <- glue('{data_dir}/centerlines/10_centerlines/{checked_site_yr}')
# fs::dir_create(lines_dir)
  
site_sf_list <- fs::dir_ls(lines_dir, glob = glue('*{checked_site_yr}-{my_group_id}-riverline*.shp')) %>% purrr::map(~st_read(.x)) 
site_sf <- site_sf_list %>% bind_rows()
site_sf <- site_sf %>% 
  dplyr::mutate(length_km = as.numeric(st_length(site_sf)/1000)) %>% 
  dplyr::mutate(length_km = round(length_km, 2)) %>%
  dplyr::select(site_yr_id, group_id, seg_id, length_km)
site_sf %>% st_write(glue('{lines_dir}/{checked_site_yr}-{my_group_id}-riverlines.shp'), append = FALSE)

site_sf %>% st_union() %>% st_length()/1000
site_sf$length_km %>% sum()
```

```{r}
site_sf %>% 
  ggplot() + 
  geom_sf(data = my_riv_fill10, fill = 'dodgerblue', col = 'dodgerblue') +
  # geom_sf(data = my_riv_fill10, aes(fill = factor(rand_id), col = factor(rand_id))) +
  # scale_fill_viridis_d() + scale_color_viridis_d() +
  geom_sf(col = 'red') + 
  theme_void()

ggsave(glue('{checked_site_yr}-{my_group_id}.pdf'), width = 12)

```

```{r}
library(leaflet)
line4326 <- st_transform(site_sf, 'EPSG:4326')

leaflet() %>%
  addTiles(group = 'bg') %>%
  addProviderTiles(providers$Esri.WorldImagery, group = 'esri') %>%
  addPolygons(data = riv4326, color = 'cyan', group = 'riv',
              fillOpacity = 1, opacity = 1, label = ~glue('{rand_id}')) %>%
  addPolylines(data = line4326, col = 'red', opacity = 1, group = 'line') %>%
  addLayersControl(overlayGroups = c('bg','riv', 'esri', 'line'))
```

## end!


```{r}
glue('{data_dir}/centerlines/10_centerlines') %>% fs::dir_ls() %>% basename()
```

```{r}
dir_ls(glue('{data_dir}/centerlines/10_centerlines'), glob = glue('*-riverlines.shp'), recurse = TRUE) %>% 
  purrr::map(~st_read(.x)) %>%
  purrr::map(~st_union(.x)) %>% 
  purrr::map(~st_length(.x)/1000)

# site_sf %>% st_union() %>% st_length()/1000
# site_sf$length_km %>% sum()
```

